Lesson 5
========================================================
getwd()
setwd('/Users/Jack/Dropbox/R/udacity-github/EDA_Course_Materials/lesson5')
library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv', sep='\t')

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + 
  geom_boxplot() +
  stat_summary(fun.y='mean', geom='point', shape=5)


ggplot(aes(x=age, y=friend_count), data=subset(pf, !is.na(gender))) + 
  geom_line(aes(color=gender), stat='summary', fun.y='median')

### 
library(dplyr)
### chain function together, create new data frame pf.fc_by_age_gender

pf.fc_by_age_gender <- pf %>%
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count=mean(friend_count),
            median_friend_count=quantile(friend_count, .5),
            n=n()) %>%
  ungroup() %>%
  arrange(age)

head(pf.fc_by_age_gender)

```




***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
ggplot(aes(x=age, y=median_friend_count), data=pf.fc_by_age_gender) + 
  geom_line(aes(color=gender))
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
#install.packages('reshape2')
library(reshape2)
# original data
head(pf.fc_by_age_gender)

# change to wide with limited column
pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender, age ~ gender, value.var = 'median_friend_count' )
head(pf.fc_by_age_gender.wide)

#change to long
pf.fc_by_age_gender.long <- melt(pf.fc_by_age_gender.wide, id.vars = c("age"))
head(pf.fc_by_age_gender.long,10)
tail(pf.fc_by_age_gender.long,10)
```

***

### Ratio Plot
Notes:

```{r Ratio Plot}
library(ggplot2)
ggplot(aes(x=age, y=female/male), data=pf.fc_by_age_gender.wide) +
  geom_line() +
  geom_hline(yintercept=1, alpha=0.3, linetype=2)

```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
#create variable year_joined and assign it to dataframe to pf
# that means year_joined will show up on your pf data frame
pf$year_joined <- 2014 - floor(pf$tenure/365)
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
summary(pf$year_joined)
table(pf$year_joined)

?cut
pf$year_joined.bucket <-cut(pf$year_joined, c(2004, 2009, 2011, 2012, 2014))
head(pf$year_joined.bucket)
table(pf$year_joined.bucket, useNA='ifany')
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
ggplot(aes(x=age, y=friend_count), data=subset(pf, !is.na(year_joined.bucket)))+
  geom_line(aes(color=year_joined.bucket),
            stat='summary', fun.y=median) 


```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}

ggplot(aes(x=age, y=friend_count), data=subset(pf, !is.na(year_joined.bucket)))+
  geom_line(aes(color=year_joined.bucket),
            stat='summary', fun.y=mean) +
  geom_line(stat='summary', linetype=2, fun.y=mean)
```

***

### Friending Rate
Notes:

```{r Friending Rate}
with(subset(pf, tenure>=1), summary(friend_count/tenure))
```

***

### Friendships Initiated
Notes:

What is the median friend rate? 0.22
 
What is the maximum friend rate? 439

```{r Friendships Initiated}
ggplot(aes(x=tenure, y=friendships_initiated/tenure),
       data=subset(pf, tenure>=1)) +
  geom_line(aes(color=year_joined.bucket))
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

p1 <- ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

p2 <- ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

p3 <- ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

p4 <-ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)



p5 <-ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure > 1)) +
  geom_smooth(aes(color=year_joined.bucket))


library("gridExtra")
grid.arrange(p1,p2,p3,p4,p5, ncol=1)

```
***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes: yogurt.csv
```{r load the Yogurt data}
getwd()
yo <-read.csv('yogurt.csv')
str(yo)

#chang the id from an int to a factor variable
yo$id <- factor(yo$id)
str(yo)
```
***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
qplot(data=yo, x=price, fill=I('#F79420'))
qplot(data=yo, x=price, fill=I('#F79420'),binwidth=13)

```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
# summary give you five number data: min, 1st quantile, 3rd, max, mean and median
summary(yo)
# table gives you the distribution of the variable
table(yo$price)
# length give you the count
length(unique(yo$price))

# transform can add new variables to the dataframe
yo <-transform(yo, all.purchase = strawberry + blueberry + pina.colada + plain + mixed.berry)
summary(yo$all.purchase)
```

***

### Prices over Time
Notes:

```{r Prices over Time}
ggplot(aes(x=time, y=price), data=yo) +
  geom_point()

#jitter makes your data a little transparent
ggplot(aes(x=time, y=price), data=yo) +
  geom_jitter(alpha=0.25, shape=21, fill=I('#F79420'))
```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
# set the seed for reproducible results
set.seed(100)
sample.ids <- sample(levels(yo$id), 8)
sample.ids

ggplot(aes(x=time, y=price), data=subset(yo, id%in% sample.ids)) +
  facet_wrap(~id) +
  geom_line() +
  geom_point(aes(size=all.purchase), pch=1)
```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

```{r Scatterplot Matrix}
#install.packages("GGally")
library(GGally)
theme_set(theme_minimal(20))

#set the seed for reproducbile results
set.seed(1836)
pf_subset <- pf[,c(2:15)]
names(pf_subset)
ggpairs(pf_subset[sample.int(nrow(pf_subset), 1000), ])

```


***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")

# change the colnames to produce a nicer plot
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

