Lesson 4
========================================================

***
getwd()
setwd('/Users/Jack/Dropbox/R/udacity-github/EDA_Course_Materials/lesson4')
getwd()
### Scatterplots and Perceived Audience Size
Notes: its better to use scatterplots to identify the relationship between two variables

***

### Scatterplots
Notes:

```{r Scatterplots}
library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv', sep='\t')

qplot(x=age, y=friend_count, data=pf)
### alternate way

qplot(age,friend_count, data=pf)

```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes: The main difference between ggplot and qplot is: ggplot needs to specify which geom and ggplot use aes wrap, ggplot come plot more complex plot
use + to add one new layer 
its better to add one layer at one time

```{r ggplot Syntax}
ggplot(aes(x=age, y=friend_count), data=pf) + geom_point()

summary(pf$age)
 
ggplot(aes(x=age, y=friend_count), data=pf) + geom_point() + xlim(13,90)
```

***

### Overplotting
Notes: set the transport using alpha param, alpha=1/20 means use 1 point to express 20 points at the same location

```{r Overplotting}
ggplot(aes(x=age, y=friend_count), data=pf) + 
  geom_point(alpha=1/20) + 
  xlim(13,90)

ggplot(aes(x=age, y=friend_count), data=pf) + 
  geom_jitter(alpha=1/20) + 
  xlim(13,90)

ggplot(aes(x=age, y=friend_count), data=pf) + 
  geom_jitter(alpha=0.05) + 
  xlim(13,90)
```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(x=age, y=friend_count), data=pf) + 
  geom_point(alpha=0.05, position = position_jitter(h=0)) + 
  xlim(13,90) +
  coord_trans( y="sqrt")
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(aes(x=age, y=friendships_initiated), data=pf) + 
  geom_point(alpha=1/10, position=position_jitter(h=0)) + 
  xlim(13,90) + 
  coord_trans(y="sqrt")
```

#### What do you notice?

***


### Conditional Means
Notes:

```{r Conditional Means}
#install.packages('dplyr')
library(dplyr)
age_groups <- group_by(pf, age)
pf.fc_by_age <- summarise(age_groups,
          friend_count_mean = mean(friend_count),
          friend_count_media = mean(friend_count),
          n = n())
pf.fc_by_age <- arrange(pf.fc_by_age, age)
head(pf.fc_by_age)
```

Conditional means alernate code, use %>% to chain functions

```{r Conditional Means Plot}
pf.fc_by_age <- pf %>%
  group_by(age) %>%
  summarise(friend_count_mean=mean(friend_count),
            friend_couint_median=median(friend_count),
            n=n()) %>%
  arrange(age)

head(pf.fc_by_age,20)

#using the new data frame "pf.fc_by_age" to plot relationship better age vs friend_count

ggplot(aes(x=age, y=friend_count_mean), data=pf.fc_by_age) + 
  geom_line() + 
  xlim(13,90) 

```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
ggplot(aes(x=age, y=friend_count), data=pf) + 
  geom_point(alpa=0.05, position=position_jitter(h=0),
            color = 'orange') +
  xlim(13,90) +
  coord_trans(y='sqrt')  +
  geom_line(stat='summary', fun.y = mean) +
  geom_line(stat='summary', fun.y=quantile, probs = .1, linetype = 2, color = 'blue')   +
  geom_line(stat='summary', fun.y=quantile, probs = .5, linetype = 2, color = 'white') +
  geom_line(stat='summary', fun.y=quantile, probs = .9, linetype =2, color='red')


```

#### use coord_cartesian
```{r}
ggplot(aes(x=age, y=friend_count), data=pf) + 
  geom_point(alpa=0.05, position=position_jitter(h=0),
            color = 'orange') +
  coord_cartesian(xlim=c(13,90), ylim=c(0,1000)) +
  geom_line(stat='summary', fun.y = mean) +
  geom_line(stat='summary', fun.y=quantile, probs = .1, linetype = 2, color = 'blue')   +
  geom_line(stat='summary', fun.y=quantile, probs = .5, linetype = 2, color = 'white') +
  geom_line(stat='summary', fun.y=quantile, probs = .9, linetype =2, color='red')

```


***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Look up the documentation for the cor.test function.
Notes:

```{r Correlation}
?cor.test()
cor.test(pf$age, pf$friend_count, method='pearson')
## or you can use
with(pf, cor.test(age, friend_count, method='pearson'))
```


### Correlation on Subsets
Notes:
```{r}
with(subset(pf, age<=70), cor.test(age, friend_count, method='pearson'))
```


***


### Correlation Methods
Notes:
```{r}
with(subset(pf, age<=70), cor.test(age, friend_count, method='spearman'))
```
***

## Create Scatterplots
Notes:

```{r}
ggplot(aes(x=www_likes_received, y=likes_received), data=pf) +
  geom_point() +
  xlim(0, quantile(pf$www_likes_received, 0.95)) +
  ylim(0, quantile(pf$likes_received, 0.95)) + 
  geom_smooth(method='lm',color='red')
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
cor.test(pf$www_likes_received, pf$likes_received, method='pearson')
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
#install.packages('alr3')
library(alr3)
data(Mitchell)
?Mitchell
head(Mitchell)

```

Create your plot!

```{r Temp vs Month}
ggplot(aes(x=Month, y=Temp), data=Mitchell) + 
  geom_point()

## or you can use qplot
qplot(data=Mitchell, Month, Temp)
cor.test(Mitchell$Month, Mitchell$Temp)
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
ggplot(aes(x=Month, y=Temp), data=Mitchell) + 
  geom_point()
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
range(Mitchell$Month)

ggplot(aes(x=Month, y=Temp), data=Mitchell) + 
  geom_point() + 
  scale_x_discrete(breaks=seq(0,203,12))
```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
ggplot(aes(x=age, y=friend_count_mean),data=pf.fc_by_age) +
  geom_point() + 
  geom_line()

head(pf.fc_by_age)
pf.fc_by_age[17:19,]
```

***

### Age with Months Means

```{r Age with Months Means}
pf$age_with_months <- pf$age + (12-pf$dob_month)/12
head(pf$age_with_months)
```

Programming Assignment
```{r Programming Assignment}
library(dplyr)
pf.fc_by_age_months <- pf %>%
  group_by(age_with_months) %>%
  summarise(friend_count_mean=mean(friend_count),
            friend_count_median=median(friend_count),
            n=n()) %>%
  arrange(age_with_months)

head(pf.fc_by_age_months)
### another way
age_with_months_groups <- group_by(pf, age_with_months)
pf.fc_by_age_months2 <- summarise(age_with_months_groups,
                          friend_count_mena=mean(friend_count),
                          friend_count_median=median(friend_count),
                          n=n())
pf.fc_by_age_months2 <- arrange(pf.fc_by_age_months2, age_with_months)



```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
ggplot(aes(x=age_with_months, y=friend_count_mean), 
       data=subset(pf.fc_by_age_months,age_with_months<71)) + 
         geom_line()
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
p1 <- ggplot(aes(x=age_with_months, y=friend_count_mean), 
       data=subset(pf.fc_by_age_months,age_with_months<71)) + 
  geom_line() + 
  geom_smooth()

p2 <- ggplot(aes(x=age, y=friend_count_mean),
       data=subset(pf.fc_by_age, age<71)) +
  geom_line() +
  geom_smooth()

p3 <- ggplot(aes(x=round(age/5)*5, y=friend_count),
             data=subset(pf, age<71)) +
  geom_line(stat='summary', fun.y='mean')

library(gridExtra)
grid.arrange(p1, p2, p3, ncol=1)
```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

