---
output: html_document
---
Lesson 3
========================================================

getwd()
setwd('/Users/Jack/Dropbox/R/udacity-github/EDA_Course_Materials/lesson3')
getwd()
### What to Do First?
Notes:
```{r}
getwd()
setwd('/Users/Jack/Dropbox/R/udacity-github/EDA_Course_Materials/lesson3')
getwd()
```



### Pseudo-Facebook User Data
Notes:

```{r Pseudo-Facebook User Data}
list.files()
pf <-read.csv('pseudo_facebook.tsv', sep='\t')
names(pf)
```

 

### Histogram of Users' Birthdays
Notes:

```{r Histogram of Users\' Birthdays}
#install.packages('ggplot2')
library(ggplot2)
#names(pf)
qplot(data=pf, x=dob_day) +
  scale_x_discrete(breaks=1:31) + 
  facet_wrap(~dob_month, ncol=3)
```
#### friends count
 
```{r}
qplot(data=pf, x=friend_count)
```
### Equivalent ggplot syntax:
```{r}
ggplot(aes(x = friend_count), data = pf) +
  geom_histogram()
```

#### What are some things that you notice about this histogram?
Response:

 

### Moira's Investigation
Notes:

 

### Estimating Your Audience Size
Notes:

 

#### Think about a time when you posted a specific message or shared a photo on Facebook. What was it?
Response:

#### How many of your friends do you think saw that post?
Response:

#### Think about what percent of your friends on Facebook see any posts or comments that you make in a month. What percent do you think that is?
Response:

 

### Perceived Audience Size
Notes:

 
### Faceting
Notes:

```{r Faceting}

```

#### Let’s take another look at our plot. What stands out to you here?
Response:

 

### Be Skeptical - Outliers and Anomalies
Notes:

 

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response:

 

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
qplot(data=pf, x=friend_count)
```

#### How is this plot similar to Moira's first plot?
Response:

 

### Limiting the Axes
Notes:

```{r Limiting the Axes}
qplot(x=friend_count, data=pf, xlim=c(0,1000))
```

### Exploring with Bin Width and Faceting Friend Count
Notes:
```{r}
qplot(x=friend_count, data=pf, binwidth=25) +
  scale_x_continuous(limits=c(0,1000), breaks = seq(0,1000,50)) +
  facet_wrap(~gender)

qplot(x=friend_count, data=pf, binwidth=25) +
  scale_x_continuous(limits=c(0,1000), breaks = seq(0,1000,50)) +
  facet_grid(gender~.)
```
 

### Omitting NA Values, The missing values
Notes:

```{r Omitting NA Values}
qplot(x=friend_count, data=subset(pf, !is.na(gender)), binwidth=10) +
  scale_x_continuous(lim=c(0,1000), breaks =seq(0,1000,50)) + 
  facet_wrap(~gender)
```

 

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
table(pf$gender)
by(pf$friend_count,pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response:

#### What's the difference between the median friend count for women and men?
Response:

#### Why would the median be a better measure than the mean?
Response:

 

### Tenure
Notes:

```{r Tenure}
qplot(x=tenure, data=pf, 
      color =I('black'), fill = I('#099DD9'))

qplot(x=tenure, data=pf, binwidth=30,
      color =I('black'), fill = I('#099DD9'))
```

 

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
qplot(x=tenure/365, data=pf, binwidth=1,
      color =I('black'), fill = I('#099DD9'))

qplot(x=tenure/365, data=pf, binwidth=0.25,
      color =I('black'), fill = I('#099DD9')) +
  scale_x_continuous(breaks=seq(1,7,1), limits=c(0,7))
```

 

### Labeling Plots
Notes:

```{r Labeling Plots}
qplot(x=tenure/365, data=pf, binwidth=0.25,
      color =I('black'), fill = I('#099DD9')) +
  scale_x_continuous(breaks=seq(1,7,1), limits=c(0,7)) +
  xlab('Number of years using Facebook') +
  ylab('Number of users in sample')
```

 

### User Ages
Notes:

```{r User Ages}
qplot(x=age, data=pf, binwidth=1,
      color =I('black'), fill = I('#099DD9')) +
    scale_x_continuous(breaks=seq(1,100,10), limits=c(0,120)) +
  xlab('age of friend users') +
  ylab('Number of users in sample')
```

#### What do you notice?
Response:

 

### The Spread of Memes
Notes:

 

### Lada's Money Bag Meme
Notes:

 

### Transforming Data
Notes:
```{r}
a <- pf$friend_count
b <- log10(pf$friend_count+1)
c <- sqrt(pf$friend_count)

summary(a)
summary(b)
summary(c)

library("ggplot2")
ap <- qplot(x=a, data=pf, binwidth=1)
bp <- qplot(x=b, data=pf, binwidth=1)
cp <- qplot(x=c, data=pf, binwidth=1)
#ap
#install.packages("gridExtra")
library("gridExtra")
grid.arrange(ap, bp, cp, ncol=3)

### alternate ways 1
p1 <- qplot(x=a, data=pf, binwidth=1)
p2 <- qplot(x=b, data=pf, binwidth=1)
p3 <- qplot(x=c, data=pf, binwidth=1) 
grid.arrange(p1, p2, p3, ncol=1)


```

 

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}

### alternate ways 3, using scales
p1 <- ggplot(aes(x=a), data=pf) + geom_histogram() + xlab("friend_count")
p2 <- p1 + scale_x_log10() + xlab("log10")
p3 <- p1 + scale_x_sqrt() + xlab("sqrt")
grid.arrange(p1, p2, p3, ncol=1)

### the difference between qplot and ggplot
### Is the labeling difference
logScale <- qplot(x=log10(friend_count),data=pf)

countScale <- ggplot(aes(x=friend_count), data=pf) +
  geom_histogram()+
  scale_x_log10()

grid.arrange(logScale, countScale, ncol=2)

```



### Frequency Polygons

```{r Frequency Polygons}
qplot(x=friend_count, data=subset(pf,!is.na(gender)),
      binwidth=10) +
  scale_x_continuous(lim=c(0,1000), breaks=seq(0,1000,50))+
  facet_wrap(~gender)

qplot(x=friend_count, y =..count../sum(..count..), 
      data=subset(pf, !is.na(gender)),
      xlab="friend connt",
      ylab="Proportion of Users with that friend  count",
      binwidth=10, geom="freqpoly", color =gender) +
  scale_x_continuous(lim=c(0,1000), breaks=seq(0,1000,50))
```

 

### Likes on the Web
Notes:

```{r Likes on the Web}
qplot(x=www_likes, data=pf, 
      xlab="www_likes",
      ylab="count",
       geom="freqpoly", color = gender) +
  scale_x_continuous() + 
  scale_x_log10()
  #scale_x_continuous()

### using ggplot
ggplot(aes(x = www_likes), data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender)) +
  scale_x_log10()

### what's the www_like count for males and females
by(pf$www_likes, pf$gender, summary)
by(pf$www_likes, pf$gender, sum)

```


 

### Box Plots
Notes:

```{r Box Plots}
qplot(x=gender, y=friend_count, 
      data=subset(pf, !is.na(gender)),
      geom='boxplot')


```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}
qplot(x=gender, y=friend_count, 
      data=subset(pf, !is.na(gender)),
      geom='boxplot', ylim=c(0,1000))

qplot(x=gender, y=friend_count, 
      data=subset(pf, !is.na(gender)),
      geom='boxplot') +
  scale_y_continuous(limits=c(0,1000))

qplot(x=gender, y=friend_count, 
      data=subset(pf, !is.na(gender)),
      geom='boxplot') +
  coord_cartesian(ylim=c(0,1000))

qplot(x=gender, y=friend_count, 
      data=subset(pf, !is.na(gender)),
      geom='boxplot') +
  coord_cartesian(ylim=c(0,250))



```

 

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}
by(pf$friend_count, pf$gender, summary)

```

#### On average, who initiated more friendships in our sample: men or women?
Response:
#### Write about some ways that you can verify your answer.
Response:
```{r Friend Requests by Gender}
qplot(x=gender, y=friendships_initiated,
      data=subset(pf, !is.na(gender)), geom="boxplot") + 
  coord_cartesian(ylim=c(0,150))

by(pf$friendships_initiated, pf$gender, summary)
```

Response:

 

### Getting Logical
Notes:

```{r Getting Logical}
mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes >0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)

sum(pf$mobile_check_in ==1 )/length(pf$mobile_check_in)

```

Response:

 

### Analyzing One Variable
Reflection:

 

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!